---
output:
  word_document: default
  html_document: default
---
```{r, Setup, eval = T, echo = F, message = F, warning = F}
# Set up environment
path <- "/Users/Alex/Desktop/"
path2 <- "/Users/Alex/"

library(reshape2)
library(DESeq2)
library(rain)

# Sample data
metadata <- read.csv(file = paste(path2, "Desktop/geodes/bioinformatics_workflow/R_processing/sample_metadata.csv", sep = ""), header = T)

```

```{r, Trout, echo = F, eval = T}

tnorm <- read.csv(paste(path, "geodes_data_tables/Trout_ID90_normalized_readcounts.csv", sep = ""), header = T, row.names = 1)
tnorm <- tnorm[, which(colnames(tnorm) != "GEODES065.nonrRNA")]
colnames(tnorm) <- gsub(".nonrRNA", "", colnames(tnorm))
trout_key <- read.csv(paste(path, "geodes_data_tables/Trout_ID90_genekey_geneclassifications_2018-11-28.csv", sep = ""), header = T)


```
**Table 3. Gene expression in day vs. night by functional categories in Trout Bog. ** We aggregated timepoints by day (9AM, 1PM, 5PM) and night (9PM, 1AM, 5AM) to compare differential gene expression. This analysis includes the top 20,000 most expressed genes. Functional categories were determined based on gene annotations. Genes with cyclic trends were detected using RAIN (Thaben and Westermark, 2014), while p-values of day vs. night read totals per sample were calculated using a two-tailed t-test. The day/night ratio is the sum of reads assigned to that category in day divided by the sum at night. A ratio greater than one indicates higher expression in day, while a ratio less than one indicates higher expression at night.


```{r, Trout_table, eval = T, include = F}

##### DESeq differential expression testing

abun_tnorm <- tnorm[order(rowSums(tnorm), decreasing = T), ]
abun_tnorm <- abun_tnorm[1:20000,]

colnames(abun_tnorm) <- gsub(".nonrRNA", "", colnames(abun_tnorm))

input <- as.matrix(abun_tnorm)
input <- input/500
input <- round(input, digits = 0)

conditions <- metadata$Time[match(colnames(abun_tnorm), metadata$Sample)]
conditions[which(conditions == 9 | conditions == 13 | conditions == 17)] <- "day"
conditions[which(conditions == 5 | conditions == 21 | conditions == 1)] <- "night"
coldata <- data.frame(samples = colnames(abun_tnorm), conditions)

cds <- DESeqDataSetFromMatrix(countData = input,
                              colData = coldata,
                              design = ~ conditions)

cds <- estimateSizeFactors(cds) 
cds <- estimateDispersions(cds) 


dds <- DESeq(cds)
res <- results(dds)

reskey <- trout_key[match(rownames(res), trout_key$Gene), ]

# Write a function to report trends in metabolic categories
pathway.stats <- function(searchterm){
  genes <- reskey$Gene[grep(searchterm, reskey$Product)]
  numgenes <- length(genes)
  gene_results <- res[match(genes, rownames(res)), ]
  gene_counts <- abun_tnorm[match(genes, rownames(abun_tnorm)),]
  gene_results$Day <- rowSums(gene_counts[which(conditions == "day")])
  gene_results$Night <- rowSums(gene_counts[which(conditions == "night")])
  gene_results$Leaning <- gene_results$Day > gene_results$Night
  percent_genes_day <- length(which(gene_results$padj < 0.05 & gene_results$Leaning == TRUE))/length(genes) * 100
  percent_genes_night <- length(which(gene_results$padj < 0.05 & gene_results$Leaning == FALSE))/length(genes) * 100
  percent_reads_day <- sum(gene_results$Day[which(gene_results$padj < 0.05 & gene_results$Leaning == TRUE)])/sum(gene_results$Day) * 100
  percent_reads_night <- sum(gene_results$Night[which(gene_results$padj < 0.05 & gene_results$Leaning == FALSE)])/sum(gene_results$Night) * 100
  day_totals <- colSums(gene_counts[, which(conditions == "day")])
  night_totals <- colSums(gene_counts[which(conditions == "night")])
  sig <- t.test(day_totals, night_totals, alternative = "two.sided")
  
  p.genes <- rownames(gene_results)[which(gene_results$padj < 0.05)]
  if(length(p.genes) > 0){
    p.table <- abun_tnorm[match(p.genes, rownames(abun_tnorm)),]
    p.results <- rain(t(p.table), deltat = 4, period = 24, measure.sequence = table(metadata$Timepoint[match(colnames(p.table), metadata$Sample)]), verbose = T, adjp.method = "Bonferroni")
    percent_cyclic_genes <- length(which(p.results$pVal < 0.05 & p.results$phase == 12))/length(genes) * 100
    gene_results$TotalReads <- apply(gene_results[ ,7:8], 1, sum)
    percent_cyclic_reads <- sum(gene_results$TotalReads[match(p.genes, rownames(gene_results))])/sum(gene_results$TotalReads) * 100
  }else{
    percent_cyclic_genes = 0
    percent_cyclic_reads = 0
  }
  
  report <- c(numgenes, round(percent_genes_day, 2), round(percent_reads_day, 2), round(percent_genes_night, 2), round(percent_reads_night, 2), round(percent_cyclic_genes, 2), round(percent_cyclic_reads, 2), mean(day_totals), mean(night_totals), round(sig$p.value, 2))
  names(report) <- c("%DayGenes", "%DayReadsContributed", "%NightGenes", "%NightReadsContributed", "%CyclicGenes", "%CyclicReadsContributed", "DayReads", "NightReads", "t-test")
  return(report)
}

# Run the function

photo <- pathway.stats("photo|Photo")
rhodo <- pathway.stats("opsin|Opsin")
rubisco <- pathway.stats("rbcL|ribulose-bisphosphate carboxylase")
polyamines <- pathway.stats("putrescine|Putrescine|spermidine|Spermidine")
rTCA <- pathway.stats("citrate lyase|Citrate lyase")
chitinase <- pathway.stats("Chitobiase|chitobiase|chitinase|Chitinase")
ap <- pathway.stats("alkaline phosphatase|Alkaline phosphatase")
cellulase <- pathway.stats("cellulase|cellulose")
ros <- pathway.stats("peroxidase|peroxide|catalase")
nitrite_reduction <- pathway.stats("nitrite reductase|Nitrite reductase|nitrite oxidoreductase|Nitrite oxidoreductase")
protease <- pathway.stats("protease")
carbox <- pathway.stats("carboxylate transport")
ribose <- pathway.stats("ribose transport")
gen.sugars <- pathway.stats("sugar transport|carbohydrate ABC transport|Carbohydrate-selective porin")
raff <- pathway.stats("raffinose/stachyose/melibiose transport")
gluc <- pathway.stats("glucose/mannose transport")
rham <- pathway.stats("rhamnose transport")
xyl <- pathway.stats("xylose transport")
fruct <- pathway.stats("fructose transport")
gh <- pathway.stats("glycoside hydrolase|Glycoside hydrolase|glycosyl hydrolase")
meth <- pathway.stats("ammonia monooxygenase|methane monoxygenase")
urea <- pathway.stats("urease")
amino <- pathway.stats("amino acid transport")
```


|                         | Number of genes | % Genes more expressed in day      | % Genes more expressed at night   | % Cyclic genes (12 hr phase)  | p-value from t-test of day vs. night read totals | Day/night ratio |
| -----------------------|:-----------------:|:------------------------:|:------------------------:|:------------------------:|:------------------------:|:-----:|
| Photosynthesis | `r round(photo[1], 2)` | `r round(photo[2], 2)` | `r round(photo[4], 2)` | `r round(photo[6], 2)` | `r round(photo[10], 2)` | `r round(photo[8]/photo[9], 2)` |
| RuBisCO | `r round(rubisco[1], 2)` | `r round(rubisco[2], 2)` | `r round(rubisco[4], 2)` | `r round(rubisco[6], 2)` | `r round(rubisco[10], 2)` | `r round(photo[8]/photo[9], 2)` |
| Glycoside hydrolases | `r round(gh[1], 2)` | `r round(gh[2], 2)` | `r round(gh[4], 2)` | `r round(gh[6], 2)` | `r round(gh[10], 2)` | `r round(gh[8]/gh[9], 2)` |
| Polyamines | `r round(polyamines[1], 2)` | `r round(polyamines[2], 2)` | `r round(polyamines[4], 2)` | `r round(polyamines[6], 2)` | `r round(polyamines[10], 2)` | `r round(polyamines[8]/polyamines[9], 2)` |
| Reactive oxygen species | `r round(ros[1], 2)` | `r round(ros[2], 2)` | `r round(ros[4], 2)` | `r round(ros[6], 2)` | `r round(ros[10], 2)` | `r round(ros[8]/ros[9], 2)` |
| Protease | `r round(protease[1], 2)` | `r round(protease[2], 2)` | `r round(protease[4], 2)` | `r round(protease[6], 2)` | `r round(protease[10], 2)` | `r round(protease[8]/protease[9], 2)` |
| Ribose transport | `r round(ribose[1], 2)` | `r round(ribose[2], 2)` | `r round(ribose[4], 2)` | `r round(ribose[6], 2)` | `r round(ribose[10], 2)` | `r round(ribose[8]/ribose[9], 2)` |
| General sugar transport | `r round(gen.sugars[1], 2)` | `r round(gen.sugars[2], 2)` | `r round(gen.sugars[4], 2)` | `r round(gen.sugars[6], 2)` | `r round(gen.sugars[10], 2)` | `r round(gen.sugars[8]/gen.sugars[9], 2)` |
| Xylose transport | `r round(xyl[1], 2)` | `r round(xyl[2], 2)` | `r round(xyl[4], 2)` | `r round(xyl[6], 2)` | `r round(xyl[10], 2)` | `r round(xyl[8]/xyl[9], 2)` |
| Methane/ammonia monooxygenase | `r round(meth[1], 2)` | `r round(meth[2], 2)` | `r round(meth[4], 2)` |  `r round(meth[6], 2)` | `r round(meth[10], 2)` | `r round(meth[8]/meth[9])` |
| Amino acid transport | `r round(amino[1], 2)` | `r round(amino[2], 2)` | `r round(amino[4], 2)` | `r round(amino[6], 2)` | `r round(amino[10], 2)` | `r round(amino[8]/amino[9], 2)` |
