# Mapping

####Goal of this analysis

Who's active in our samples? What genes are being expressed? We could try to classify and annotate each read in the metatranscriptomes, but that wouldn't be very accurate. Instead we're going to map the reads to a database of freshwater genomes to get the classification and  annotation of every read we can.

##### Approach

The input files are the output from the workflow in 01rRNA_removal. In that workflow, I took quality filter reads produced by the JGI and ran them through sortmerna to get only nonrRNA reads to map. The database I'm using is genomes assembled from metagenomes (MAGs) and single amplified genomes (SAGs) from the same lakes as my metatranscriptomes. It's been curated by both the JGI and our lab, and contains additional genomes generated from the metagenomes and single cell preservations we collected during the GEODES field campaign. We're hoping that we'll get the best mapping possible by using genomes from the same study sites. We're also mapping to an internal standard that we added during extraction to approximate the absolute read counts in our metatranscriptomes. I'm going to perform this mapping competitively using the program bwa, then summarize the results using samtools and htseq.

## Most recent workflow. 
####Use this if you want to replicate our protocol. Updated ...


#Lab Notebook

####2017-02-13

There's quite a few programs involved in this step. I'm using:

- bwa (Burrows-Wheeler Aligner, v0.7.12), https://sourceforge.net/projects/bio-bwa/files/ 
- samtools (v1.3.1), https://sourceforge.net/projects/samtools/files/
- htseq (v0.6.1), http://www-huber.embl.de/users/anders/HTSeq/doc/install.html#install
- python (gzipped source tarball v2.7.13), https://www.python.org/downloads/release/python-2713/

The general plan is to:
- copy in the nonrRNA file of interest
- run bwa
- convert the output from .sam to .bam and index
- count reads by feature with htseq
- calculate RPKM and summarize results with a python script
- save files back to gluster

I'm basing this on Josh's script from OMD-TOIL found here https://github.com/alexlinz/OMD-TOILv2/tree/master/scripts/10_mapping/10c_competitive

But the installation on CHTC is going to be a little tricky. First I need to build my own python tarball in interactive mode. I'm following the instructions here http://chtc.cs.wisc.edu/python-jobs.shtml

First I need to start an interactive CHTC session.
interactive.sub:
```{r, eval = F}
universe = vanilla
# Name the log file:
log = interactive.log

# Name the files where standard output and error should be saved:
output = process.out
error = process.err

# If you wish to compile code, you'll need the below lines. 
#  Otherwise, LEAVE THEM OUT if you just want to interactively test!
+IsBuildJob = true
requirements = (OpSysAndVer =?= "SL6") && ( IsBuildSlot == true )

# Indicate all files that need to go into the interactive job session,
#  including any tar files that you prepared:
transfer_input_files = Python-2.7.13.tgz

# It's still important to request enough computing resources. The below 
#  values are a good starting point, but consider your file sizes for an
#  estimate of "disk" and use any other information you might have
#  for "memory" and/or "cpus".
request_cpus = 1
request_memory = 1GB
request_disk = 1GB

queue
```

Start the session with:
```{r, eval = F}
condor_submit -i interactive.sub
```

Here's what I'm typing in the interactive session:
```{r, eval = F}
mkdir python
tar -xvf Python-2.7.13.tgz
cd Python-2.7.13
../configure --prefix=$(pwd)/../python
make
make install
cd ..
ls python
ls python/bin
# Maybe I can install htseq right in here?
export PATH=$(pwd)/python/bin:$PATH
wget https://bootstrap.pypa.io/get-pip.py
python get-pip.py
pip install numpy
pip install matplotlib
pip install htseq
# hey, that worked!
tar -czvf python.tar.gz python/
exit

```

Wow, I feel like a real bioinformatician! +1 to CHTC for awesome tutorials. Now I have a tarball that I can transfer around with my jobs that includes both python and HTseq.

CHTC is closing down my submit node for maintenance tomorrow, so I'm going to stop here for now. I'll possibly do some testing on Zissou and will download my installation packages for that, but I'm not sure how two different versions of python will agree on Zissou.


####2017-02-15

We're back up and running! Today I'm going to write a script to map one metatranscriptome. I'm going to transfer all of the installs over, but examine the output and slowly add lines to the bash script as I go.

First, get sample names and shorten to one metatranscriptome.

```{r, eval = F}
for file in /mnt/gluster/amlinz/GEODES_nonrRNA_concat/*; do sample=$(basename $file |cut -d'.' -f1); echo $sample;done > samplenames.txt
head -1 samplenames.txt > test_samplenames.txt; mv test_samplenames.txt samplenames.txt
```

Here's my first stab at the submit file 02mapping.sub:
```{r, eval = F}
# 02mapping.sub
#
#
# Specify the HTCondor Universe
universe = vanilla
log = 02mapping_$(Cluster).log
error = 02mapping_$(Cluster)_$(Process).err
#
# Specify your executable, arguments, and a file for HTCondor to store standard
#  output.
executable = 02mapping.sh
arguments = $(samplename)
output = 02mapping_$(Cluster).out
#
# Specify that HTCondor should transfer files to and from the
#  computer where each job runs.
should_transfer_files = YES
when_to_transfer_output = ON_EXIT
transfer_input_files = python.tar.gz,bwa.tar.gz,samtools.tar.gz, mapping_database.tar.gz
#transfer_output_files =
#
# Tell HTCondor what amount of compute resources
#  each job will need on the computer where it runs.
Requirements = (Target.HasGluster == true)
request_cpus = 1
request_memory = 2GB
request_disk = 5GB
#
# Tell HTCondor to run every fastq file in the provided list:
queue samplename from samplenames.txt
```

I also need to build my reference genome database. Which means I need to go find and upload the internal standard fasta and gff. It also looks like the size is pretty small (< 1 GB) so I'll zip up this database after building and send it via the submit file

Here's my database building:
```{r, eval = F}
cat /mnt/gluster/amlinz/ref_genomes/fasta_files/*.fna > mapping_database.fna
cat /mnt/gluster/amlinz/ref_genomes/gff_files/*.gff > mapping_database.gff
tar -czvf mapping_database.tar.gz mapping_database.fna mapping_database.gff
rm mapping_database.gff
rm mapping_database.fna
```

Cool! I'm going to retroactively add that to the submit file above. Now for the executable. I'm going to be testing unzipping things so I know what to add to my path variable outside of this code.

Gah. Looks like bwa and samtools also need to be installed via interactive session.

####2017-02-16

Yep, more interactive installs time. Let's do this for bwa first. I'm modifying my interactive.sub to tranfer bwa instead of python.

Here's what I'm typing to install. Following instructions from the top of the README.md file in the bwa tarball, since I can't find installation instructions anywhere on their website...
```{r, eval = F}
condor_submit -i interactive.sub
# Wait for job to start
tar xvfj bwa-0.7.12.tar.bz2
cd bwa-0.7.12
make
cd ..
ls bwa-0.7.12
# bwa executable is in there!
tar czvf bwa.tar.gz bwa-0.7.12/
ls
exit

```

Now do the same thing for samtools. Installation instructions here: http://www.htslib.org/download/
I'm modifying my interactive.sub to transfer the samtools tarball this time.

```{r, eval = F}
condor_submit -i interactive.sub
# Wait for job to start
tar xvfj samtools-1.3.1.tar.bz2
cd samtools-1.3.1
make
make prefix=../samtools install
cd ..
ls samtools
#I've got a nice bin file in there now!
tar czvf samtools.tar.gz samtools/
ls
exit

```

Time to make my first executable. This is the homerun attempt -  I'm expecting lots and lots of errors.

02mapping.sh:
```{r, eval = F}
##!/bin/bash
#Map metatranscriptome reads to my database of reference genomes

#Unzip programs
tar xzf python.tar.gz
tar xvf bwa.tar.gz
tar xvf samtools.tar.gz
tar xvf mapping_database.gz

#Transfer metaT from gluster
cp /mnt/gluster/amlinz/GEODES_nonrRNA_concat/$1.fastq.gz
tar xzf $1.fastq.gz

#Update the path variable
mkdir home
export PATH=$(pwd)/python/bin:$(pwd)/samtools/bin:$(pwd)/bwa:$PATH
export HOME=$(pwd)/home

#Index the reference database
bwa index mapping_database.fna

#Run the mapping step - using 5 processors
bwa mem -t 5 mapping_database $1.fastq > $1.sam

#Manipulate the output
samtools view -b -S -o $1.bam $1.sam
samtools sort -o $1.sorted.bam -O bam -T /temp/$1 $1.bam
samtools index $1.sorted.bam
rm $1.sam
rm $1.bam

#Count reads
htseq-count -f bam -r pos -s no -a 0 -t feature -i locus_tag -m intersection-strict -o $1.feature.sam $1.sorted.bam mapping_database.gff > $1.feature.out

#Save output to gluster. Using cp/rm instead of mv so it will overwrite old output in gluster.
cp $1.feature.out /mnt/gluster/amlinz/GEODES_mapping_results/
rm $1.feature.out

#Clean up
rm -rf python
rm -rf bwa
rm -rf samtools
rm $1.fastq
rm *.sam
rm *.bam
rm mapping_database*
rm *.tar.gz
  
```

and the submit file:
02mapping.sub
```{r, eval = F}
# 02mapping.sub
#
#
# Specify the HTCondor Universe
universe = vanilla
log = 02mapping_$(Cluster).log
error = 02mapping_$(Cluster)_$(Process).err
#
# Specify your executable, arguments, and a file for HTCondor to store standard
#  output.
executable = 02mapping.sh
arguments = $(samplename)
output = 02mapping_$(Cluster).out
#
# Specify that HTCondor should transfer files to and from the
#  computer where each job runs.
should_transfer_files = YES
when_to_transfer_output = ON_EXIT
transfer_input_files = python.tar.gz,bwa.tar.gz,samtools.tar.gz, mapping_database.tar.gz
#transfer_output_files =
#
# Tell HTCondor what amount of compute resources
#  each job will need on the computer where it runs.
Requirements = (Target.HasGluster == true)
request_cpus = 5
request_memory = 5GB
request_disk = 8GB
#
# Tell HTCondor to run every fastq file in the provided list:
queue samplename from samplenames.txt
```

Here goes nothing! I'm going to list errors here and revise the scripts below.

1. Job held. Events 007 and 012 - condor_shadow rejected my job (Exec format error), and the job was placed in holding. Google suggests a problem with line endings - will use dos2unix on both the sub and sh and try again.
2. Still didn't work. The other suggested issue could be with the !/bin/bash statement. Oh look, I've got an extra #. Removing and trying again. THAT WORKED!
3. Several errors with a couple main causes. One, can't find GEODES001_nonrRNA.fastq.gz. Two, can't find bwa in the path. Three, can't find python. Four, can't find mapping_database.gz. Yikes. Solution 1, added ./ to the cp command so it knows where to put that. Solution 2, call bwa directly instead of via PATH. Solution 4, fix the typo in mapping_database.gz to mapping_database.tar.gz. Solution 3, add the requirements line to the sub script from CHTC's python instructions. Sidenote 1, the python tarball is > 50MB so transferring from gluster instead of via the submit file. Sidenote 2, increasing my disk usage space to 10GB, reducing memory to 3 GB and processors to 3. Still, only 12 computers meet these requirements, so I may need to split the job up into several steps to cut down on disk space.
4. "This does not look like a tar archive" and "cannot find GEODES001_nonrRNA.fastq.gz". Changing tar extraction to gzip extraction. Still can't find python -adding "python" before I call htseq-count.
5. Progress! the bwa index worked, but then bwa mem couldn't find the index. Python got found this time but then couldn't find htseq-count. But hey, at least things are running? And the diskspace is more reasonable - dropping down to 6. I'll try calling the database with the .fna extension, and calling htseq-count directly.
6. Well, that took an hour and a half but made it to the samtools step. Error was "failed to open file /temp/GEODES001_nonrRNA.0000.bam". I'll take the temp folder designation out. I'm not sure what that's doing anyway.Also going to set samtools to use three threads, and going to save a copy of the mapping output to my home folder so I can rerun from after the mapping step next time. BTW, only 27 computers match my current job.
7. Ooh, that's interesting. Error occured due to failure to parse line 18699 of mapping_database.gff, specifically in the attribute column. Investigating with:
```{r, eval = F}
sed -n '18699{p;q}' mapping_database.gff
```
Only thing different is that it has parantheses



02mapping.sh:
```{r, eval = F}
#!/bin/bash
#Map metatranscriptome reads to my database of reference genomes

#Transfer metaT from gluster
cp /mnt/gluster/amlinz/GEODES_nonrRNA_concat/$1.fastq.gz .
cp /mnt/gluster/amlinz/python.tar.gz .
gzip -d $1.fastq.gz

#Unzip programs
tar xzf python.tar.gz
tar xvf bwa.tar.gz
tar xvf samtools.tar.gz
tar xvf mapping_database.tar.gz

#Update the path variable
mkdir home
export PATH=$(pwd)/python/bin:$(pwd)/samtools/bin:$PATH
export HOME=$(pwd)/home

#Index the reference database
./bwa-0.7.12/bwa index mapping_database.fna

#Run the mapping step - using 5 processors
./bwa-0.7.12/bwa mem -t 3 mapping_database.fna $1.fastq > $1.sam
cp $1.sam /mnt/gluster/amlinz/

#Manipulate the output
samtools view -b -S -o $1.bam $1.sam
samtools sort -o $1.sorted.bam -O bam -T $1 -@ 3 $1.bam
samtools index $1.sorted.bam
rm $1.sam
rm $1.bam

#Count reads
python ./python/bin/htseq-count -f bam -r pos -s no -a 0 -t feature -i locus_tag -m intersection-strict -o $1.feature.sam $1.sorted.bam mapping_database.gff > $1.feature.out

#Save output to gluster. Using cp/rm instead of mv so it will overwrite old output in gluster.
cp $1.feature.out /mnt/gluster/amlinz/GEODES_mapping_results/
rm $1.feature.out

#Clean up
rm -rf python
rm -rf bwa
rm -rf samtools
rm $1.fastq
rm *.sam
rm *.bam
rm mapping_database*
rm *.tar.gz
  
```

02mapping.sub
```{r, eval = F}
# 02mapping.sub
#
#
# Specify the HTCondor Universe
universe = vanilla
log = 02mapping_$(Cluster).log
error = 02mapping_$(Cluster)_$(Process).err
requirements = (OpSys == "LINUX") && (OpSysMajorVer == 6)
#
# Specify your executable, arguments, and a file for HTCondor to store standard
#  output.
executable = 02mapping.sh
arguments = $(samplename)
output = 02mapping_$(Cluster).out
#
# Specify that HTCondor should transfer files to and from the
#  computer where each job runs.
should_transfer_files = YES
when_to_transfer_output = ON_EXIT
transfer_input_files = bwa.tar.gz,samtools.tar.gz, mapping_database.tar.gz
#transfer_output_files =
#
# Tell HTCondor what amount of compute resources
#  each job will need on the computer where it runs.
Requirements = (Target.HasGluster == true)
request_cpus = 3
request_memory = 3GB
request_disk = 10GB
#
# Tell HTCondor to run every fastq file in the provided list:
queue samplename from samplenames.txt
```
